Reanalysis Step 4: Propensity Score Analysis with Twang

### 2025-08-31 reanalysis-4-propensity_scores.qmd
parallelize propensity score calcualtion

library(twang)
library(ggplot2)
library(gridExtra)
library(knitr)

@@

# CACHING CONFIGURATION
use_cache <- TRUE  # Set to FALSE to force re-computation
cache_dir <- file.path(reanalysis_data_dir, "ps_cache")
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}

@@

@@

# Initialize storage for PS results
twang_results <- list()

# Track progress
progress_df <- data.frame(
  Imputation = integer(),
  Trees_Used = integer(),
  Weight_Min = numeric(),
  Weight_Max = numeric()
)

for (i in 1:n_imputations) {
  # Check for cached result
  cache_file <- file.path(cache_dir, paste0("twang_imp_", i, ".rds"))

  if (use_cache && file.exists(cache_file)) {
    log_message(paste0("Imputation ", i, ": Loading from cache"))
    cached_result <- readRDS(cache_file)
    twang_results[[i]] <- cached_result
    weights <- cached_result$weights
  } else {
    log_message(paste0("Imputation ", i, ": Fitting twang model"))
    imp_start <- Sys.time()

@@

    # Store results
    result <- list(
      model = twang_model,
      weights = weights,
      n_trees_used = twang_model$desc$es.mean$n.trees
    )
    twang_results[[i]] <- result

@@

    imp_time <- round(as.numeric(difftime(Sys.time(), imp_start, units = "mins")), 2)
    log_message(paste0("Imputation ", i, ": Completed in ", imp_time, " minutes, saved to cache"))
  }

  # Track progress (works for both cached and newly computed)
  progress_df <- rbind(progress_df, data.frame(
    Imputation = i,
    Trees_Used = twang_results[[i]]$n_trees_used,
    Weight_Min = round(min(weights), 3),
    Weight_Max = round(max(weights), 3)
  ))

  # Log progress with time estimates
  log_progress(i, n_imputations, "Imputation")
}

# Display progress summary
kable(progress_df, caption = "Twang Model Fitting Progress")
```

>>>>

library(twang)
library(ggplot2)
library(gridExtra)
library(knitr)
library(parallel)
library(doParallel)

@@

# CACHING CONFIGURATION
use_cache <- FALSE  # Set to FALSE to force re-computation
cache_dir <- file.path(reanalysis_data_dir, "ps_cache")
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}

@@

# Set up parallel processing
n_cores <- detectCores() - 1  # Use all cores except one
log_message(paste0("Parallel processing: Using ", n_cores, " cores"))

@@

# Function to process a single imputation
process_imputation <- function(i) {
  # Check for cached result
  cache_file <- file.path(cache_dir, paste0("twang_imp_", i, ".rds"))

  if (use_cache && file.exists(cache_file)) {
    cached_result <- readRDS(cache_file)
    weights <- cached_result$weights

    return(list(
      imputation = i,
      result = cached_result,
      trees_used = cached_result$n_trees_used,
      weight_min = round(min(weights), 3),
      weight_max = round(max(weights), 3),
      cached = TRUE,
      time_mins = 0
    ))
  } else {
    imp_start <- Sys.time()

@@

    # Store results
    result <- list(
      model = twang_model,
      weights = weights,
      n_trees_used = twang_model$desc$es.mean$n.trees
    )

@@

    imp_time <- round(as.numeric(difftime(Sys.time(), imp_start, units = "mins")), 2)

    return(list(
      imputation = i,
      result = result,
      trees_used = result$n_trees_used,
      weight_min = round(min(weights), 3),
      weight_max = round(max(weights), 3),
      cached = FALSE,
      time_mins = imp_time
    ))
  }
}

# Run parallel processing
log_message(paste0("Starting parallel processing of ", n_imputations, " imputations using ", n_cores, " cores"))
parallel_start <- Sys.time()

# Use mclapply on Unix/Mac, parLapply on Windows
if (.Platform$OS.type == "unix") {
  # Unix/Mac - use mclapply
  parallel_results <- mclapply(
    1:n_imputations,
    process_imputation,
    mc.cores = n_cores
  )
} else {
  # Windows - use parLapply
  cl <- makeCluster(n_cores)

  # Export necessary objects to cluster
  clusterExport(cl, c(
    "cache_dir", "use_cache", "ps_formula", "imputed_datasets",
    "n_trees", "ps", "get.weights"
  ))

  # Load required packages on each worker
  clusterEvalQ(cl, {
    library(twang)
  })

  parallel_results <- parLapply(cl, 1:n_imputations, process_imputation)

  stopCluster(cl)
}

parallel_time <- round(as.numeric(difftime(Sys.time(), parallel_start, units = "mins")), 2)
log_message(paste0("Parallel processing completed in ", parallel_time, " minutes"))

# Extract results and build progress dataframe
twang_results <- list()
progress_df <- data.frame(
  Imputation = integer(),
  Trees_Used = integer(),
  Weight_Min = numeric(),
  Weight_Max = numeric(),
  Cached = logical(),
  Time_Minutes = numeric()
)

for (res in parallel_results) {
  twang_results[[res$imputation]] <- res$result

  progress_df <- rbind(progress_df, data.frame(
    Imputation = res$imputation,
    Trees_Used = res$trees_used,
    Weight_Min = res$weight_min,
    Weight_Max = res$weight_max,
    Cached = res$cached,
    Time_Minutes = res$time_mins
  ))

  # Log individual result
  if (res$cached) {
    log_message(paste0("Imputation ", res$imputation, ": Loaded from cache"))
  } else {
    log_message(paste0("Imputation ", res$imputation, ": Fitted in ", res$time_mins, " minutes"))
  }
}

# Sort progress_df by imputation number
progress_df <- progress_df[order(progress_df$Imputation), ]

# Display progress summary with caching and timing information
kable(progress_df, 
      caption = paste0("Twang Model Fitting Progress (Total time: ", parallel_time, " minutes using ", n_cores, " cores)"),
      col.names = c("Imputation", "Trees Used", "Min Weight", "Max Weight", "From Cache", "Time (min)"))
```
